Covariant Phase Space via MultiSymplectic Geometry

### 2017-05-17 cpsviams.tex
Implemented submodules
Almost finished..

\usepackage{amsmath}
\usepackage[multisym, geomec,basic]{../Math-Symbols-List/commonmathsymb}
\usepackage[subpreambles=true]{standalone}
\usepackage{graphicx}
\usepackage{wrapfig}

\input{../Latex-Theorem/TheoremTemplateToninus.tex}

@@

The smooth bundle $(F:\rightarrow M)$ to be considered depends on the dynamical framework we are employing:

\begin{center}
\begin{tabular}{|c|c|c|}
	\hline
	 & Lagrangian picture & Hamiltonian picture \\
	\hline
	$F$		&	$E$		&	$\LinDualJet$	\\
	$\downarrow$ & $\downarrow$ & $\downarrow$ \\
	$M$ & $M$ & $M$ \\
	    & \emph{configuration bundle} & \emph{multiphase bundle} \\
	\hline
\end{tabular}
\end{center}

@@

The key point of the latter definition is that the \emph{Euler-Lagrange} map and the \emph{DeDonder-Weyl} map, being bundle-morphism, can be regarded as mapping from $\Conf$ to the section of a suitable vector bundle.
They have to be composed with the jet functor $j^1$.

\begin{equation}
	D_\Lagrangian[\; ] = D_\Lagrangian \circ j^2 : \quad \Gamma^\infty(E) \ni \varphi \quad \mapsto \quad D_\Lagrangian[j^2 \varphi ] \in \Gamma^\infty \left( \phi^\ast \left(V^\TwistedLinear E \right)\right)
\end{equation}

\begin{equation}
	D_\Hamiltonian[\; ] = D_\Hamiltonian \circ j^1 : \quad \Gamma^\infty(F) \ni (\varphi,\pi) \quad \mapsto \quad D_\Hamiltonian[j^1 (\varphi,\pi) ] \in \Gamma^\infty \left( (\varphi,\pi)^\ast \left(V^\TwistedLinear F \right)\right)
\end{equation}

@@

\begin{notationfix}
Noticing the recurring scheme we will make use of the following notation:
	\begin{displaymath}
		W = V^\TwistedLinear F \; ; \; V = \Gamma^\infty(W,F) \quad \Rightarrow \quad D[\phi] \in V \quad \forall \phi \in \Conf
	\end{displaymath}
\end{notationfix}

@@

\begin{wrapfigure}{L}{0.50\textwidth}
\centering
\includegraphics[width=0.49\textwidth]{Pictures/variations.png}
%\caption{\label{fig:variations}Covariant tangent fields as regular variations.}
\end{wrapfigure}

This condition stems from the fact that we want to interpret regular curves on $\Conf$ as smooth variations of field configurations. \\
Therefore, a parametrized regular curve as to be regarded as a family of diffeomerphisms, the flow generated by a vertical tangent field on $F$.\\
The pull-back are there to remark that $T_\phi \Conf$ consists of the infinitesimal smooth variations of $\phi$ thus are vertical field of $F$ \underline{along} the section $\phi$.

According to this interpretation, $T_\phi\Sol$ should consits of the generators of those variations that maps, at least infinitesimally, a solution $\phi\in\Sol$ in another solution.
A rigourosly, this can be expressed in the following definition:

@@

	\begin{definition}[Jacobi operator pertaining to $D_\cdot$ around a solution $\phi \in \Sol$]\label{Def:JacobiOp}
		\begin{displaymath}
			J[\phi] \; : \; T_\phi \Conf \rightarrow	\phi^\ast(V^\TwistedLinear F) \quad s.t. \quad \left\langle D[\phi] \right.\left| \delta\phi \right\rangle
			= \sigma \left( \left.\dfrac{\textrm{d}}{\textrm{d} \lambda} D_\cdot [\phi_\lambda]\right\rvert_{\lambda=0} \right)
		\end{displaymath}
	\end{definition}

\begin{wrapfigure}{L}{0.41\textwidth}
\centering
\includegraphics[width=0.40\textwidth]{Pictures/Jacobi.png}
\caption{\footnotesize \label{fig:jacobi}$J[\phi]$ and $D$ target on the same space.}
\end{wrapfigure}
where:
\begin{itemize}
	\item $\phi_\lambda$ is an arbitrary variation of $\phi_0 = \phi$ generated by $\delta\phi$.
	\item exactly as noted for the tangent vector of $\Conf$, we have for any $x\in M$:
		\begin{displaymath}
			\left.\dfrac{\textrm{d}}{\textrm{d} \lambda} D_\cdot [\phi_\lambda]\right\rvert_{\lambda=0}(x) \in V_M W
			\quad \Rightarrow \quad
			\left.\dfrac{\textrm{d}}{\textrm{d} \lambda} D_\cdot [\phi_\lambda]\right\rvert_{\lambda=0} \in D[\phi]^\ast \left( V_M \left( V^\TwistedLinear \right)\right)
		\end{displaymath}
		because $D_\cdot[\phi_\lambda]$ can be regarded as a section of $V^\TwistedLinear F$ over $M$, 
		thus the latter definition does not depend from the particular choice of variation $\phi_\lambda$ 
		but only from its infinitesimal generator $\delta\phi$.
	\item $\sigma$ is the canonical bundle-morphism
		\begin{displaymath}
			\sigma:  0^\ast \left( V_M W \right)  \; \rightarrow \;  0^\ast \left(V_E W \right) \simeq W
		\end{displaymath}
		 where $V_M W$ and $V_F W$ are the vertical sub-bundle of $T W$ w.r.t. $M$ and $F$ respectively 
		 and $0$ is the zero-section of $W$ over $F$.
\end{itemize}

@@

@@

\begin{wrapfigure}{L}{0.41\textwidth}
\centering
\includegraphics[width=0.40\textwidth]{Pictures/sigma.png}
  %\caption{.}
\end{wrapfigure}
Visually this is a map which "verticalize" w.r.t. $E$ the vertical vectors w.r.t. $M$ along $0(E) \subset W$

@@

\newpage
\section{Covariant approach: pre-symplectic form}
\begin{wrapfigure}{L}{0.41\textwidth}
\centering
\includegraphics[width=0.40\textwidth]{Pictures/symform.png}
  \caption{(temp)}
\end{wrapfigure}

@@

Rigourosly, we define $\forall \phi \in \Conf$, $\forall \Sigma \subset M$ $m-1$ dimensional, compact submanifold

\begin{definition}[(covariant) pre-symplectic form in $\phi$ pertaining to $\Sigma$]
	\begin{displaymath}
		\Omega_\Gamma ( \phi ) : \textrm{T}_\phi \Conf \times \textrm{T}_\phi \Conf \rightarrow \Real
\end{displaymath}		
	S.t:
	\begin{align*}
		\left\langle	\Omega_\Sigma ( \varphi) \right.\left\lvert \delta \varphi_1 , \delta \varphi_2 \right\rangle &=		
		\int_\Gamma (j^1 \varphi)^\ast \left[ i_{\delta\varphi_2} i_{\delta \varphi_1} \omega_\Lagrangian \right]
		& \qquad \textrm{(Lagrangian framework)} \\
		\left\langle	\Omega_\Sigma ( \varphi, \pi) \right.\left\lvert ( \delta \varphi_1, \delta \pi_1),
		( \delta \varphi_2, \delta \pi_2) \right\rangle &=		
		\int_\Gamma (\varphi, \pi)^\ast \left[ i_{( \delta \varphi_2, \delta \pi_2)} i_{( \delta \varphi_1, \delta \pi_1)} \omega_\Hamiltonian \right]
		& \qquad \textrm{(Hamiltonian framework)} 		
	\end{align*}
\end{definition}

We can easily recognize that at this level $J^1 E$ and $\LinDualJet$ takes a similar role

\paragraph{Outline}
The remainder of this article is organized as follows.
Section~\ref{previous work} gives account of previous work.
Our new and exciting results are described in Section~\ref{results}.
Finally, Section~\ref{conclusions} gives the conclusions.

>>>>

\usepackage{amsmath}
\usepackage[multisym, geomec,basic, diffgeo]{./Math-Symbols-List/toninus-math-symbols}
\usepackage[subpreambles=true]{standalone}
\usepackage{graphicx}
\usepackage{wrapfig}

\usepackage{Latex-Theorem/theoremtemplate}

@@

The smooth bundle to be considered depends on the dynamical framework we are employing.
We introduce the following notation, in order to treat both formalism at the same time:
\begin{notation}
	\begin{center}
		\begin{tabular}{|c|c|c|}
			\hline
			 & Lagrangian picture & Hamiltonian picture \\
			\hline
			$F$		&	$E$		&	$\LinDualJet$	\\
			$\downarrow$ & $\downarrow$ & $\downarrow$ \\
			$M$ & $M$ & $M$ \\
			    & \emph{configuration bundle} & \emph{multiphase bundle} \\
			\hline
		\end{tabular}
	\end{center}
\end{notation}

@@

The key point of the latter definition is that the \emph{Euler-Lagrange} map and the \emph{DeDonder-Weyl} map, being bundle-morphism, can be regarded as mapping from $\Conf$ to the section of a suitable vector bundle composing with the jet functor $j$.

\begin{equation}
	D_\Lagrangian[\; ] = D_\Lagrangian \circ j^2 : \quad \Gamma^\infty(E) \ni \varphi \quad \mapsto \quad D_\Lagrangian[ \varphi ] \in \Gamma^\infty \left( \phi^\ast \left(V^\TwistedLinear E \right)\right)
\end{equation}

\begin{equation}
	D_\Hamiltonian[\; ] = D_\Hamiltonian \circ j^1 : \quad \Gamma^\infty(F) \ni (\varphi,\pi) \quad \mapsto \quad D_\Hamiltonian[ (\varphi,\pi) ] \in \Gamma^\infty \left( (\varphi,\pi)^\ast \left(V^\TwistedLinear F \right)\right)
\end{equation}

@@

Again, we would like to employ a collective notation to denote the motion equation operators $D_\Lagrangian$ and $D_\Hamiltonian$:
\begin{notation}

	Since operators $D$ when applied to $\phi \in \Conf$ yields a twisted 1-form, 
	we denote them collectively as $D[\phi] \in \Gamma(\phi^\ast(V^\TwistedLinear F))$.
	When $D$ is evaluated on a vector field yields
	\begin{displaymath}
		\left\langle D[\phi] \right\vert \left. V \right\rangle =
		\phi^\ast \left( V \lrcorner \omega \right) 
		\begin{cases}
			(j^1 \varphi)^\ast \left( j^1V \lrcorner \omega_\Lagrangian \right) \\
			(\varphi, \pi)^\ast \left( V \lrcorner \omega_\Hamiltonian \right)
		\end{cases}              
        \in \Omega^m(M)
	\end{displaymath}

	where has been employed the following correspondences:
	\begin{center}
		\begin{tabular}{|c|c|c|}
			\hline
			 & Lagrangian picture & Hamiltonian picture \\
			\hline
			$\phi \in \Conf$		&	$j^1 \varphi$		&	$(\varphi, \pi)$	\\
			$V \in \Gamma(\phi^\ast V F )$ & $j^1 V$ & $V$ \\
			$\omega$ & $\omega_\Lagrangian$ & $\omega_\Hamiltonian$ \\
			\hline
		\end{tabular}
	\end{center}

	Noticing the recurring scheme we will make use of the following notation:
	\begin{displaymath}
		W = V^\TwistedLinear F \; ; \; V = \Gamma^\infty(\phi^\ast W)
	\end{displaymath}
	So,
	\begin{displaymath}
		D[\phi] \in V \quad \forall \phi \in \Conf
	\end{displaymath}
\end{notation}

@@

\begin{figure}
\centering
\includegraphics[width=0.49\textwidth]{Pictures/variations.png}
%\caption{\label{fig:variations}Covariant tangent fields as regular variations.}
\end{figure}

This condition stems from the fact that we want to interpret regular curves on $\Conf$ as smooth variations of field configurations. \\
Therefore, a parametrized regular curve as to be regarded as a family of diffeomerphisms, the flow generated by a vertical tangent field on $F$.\\
The pull-back are there to remark that $T_\phi \Conf$ consists of the infinitesimal smooth variations of $\phi$ thus are vertical field of $F$ \underline{along} the section $\phi$.
\begin{notation}
	\begin{center}
		\begin{tabular}{|c|c|c|}
			\hline
			 & Lagrangian picture & Hamiltonian picture \\
			\hline
			$\delta \phi \in T_\phi \Conf$		&	$\delta \varphi$		&	$(\delta \varphi, \delta \pi)$ \\
			\hline
		\end{tabular}
	\end{center}
\end{notation}

According to this interpretation, $T_\phi\Sol$ should consists of the generators of those variations that maps, at least infinitesimally, a solution $\phi\in\Sol$ in another solution.
Rigourosly, this can be expressed in the following definition:

@@

	\begin{definition}[Jacobi operator pertaining to $D_\cdot$ around a solution $\phi \in \Sol$]\label{Def:JacobiOp}
		\begin{displaymath}
			J(\phi) \; : \; T_\phi \Conf \rightarrow	\phi^\ast(V^\TwistedLinear F) 
			\quad s.t. \quad  \left(J(\phi)\right) [ \delta\phi ]
			= \sigma \left( \left.\dfrac{\textrm{d}}{\textrm{d} \lambda} D_\cdot [\phi_\lambda]\right\rvert_{\lambda=0} \right)
		\end{displaymath}
	\end{definition}

\begin{figure}
\centering
\includegraphics[width=0.40\textwidth]{Pictures/Jacobi.png}
\caption{\footnotesize \label{fig:jacobi}$ \left(J(\phi)\right) [ \delta\phi ]$ and $D$ target on the same space.}
\end{figure}
where:
\begin{itemize}
	\item $\phi_\lambda$ is an arbitrary variation of $\phi_0 = \phi$ generated by $\delta\phi$.
	\item exactly as noted for the tangent vector of $\Conf$, we have for any $x\in M$:
		\begin{displaymath}
			\left.\dfrac{\textrm{d}}{\textrm{d} \lambda} D_\cdot [\phi_\lambda]\right\rvert_{\lambda=0}(x) \in V_M W
			\quad \Rightarrow \quad
			\left.\dfrac{\textrm{d}}{\textrm{d} \lambda} D_\cdot [\phi_\lambda]\right\rvert_{\lambda=0} \in D[\phi]^\ast \left( V_M \left( V^\TwistedLinear \right)\right)
		\end{displaymath}
		because $D_\cdot[\phi_\lambda]$ can be regarded as a section of $V^\TwistedLinear F$ over $M$, 
		thus the latter definition does not depend from the particular choice of variation $\phi_\lambda$ 
		but only from its infinitesimal generator $\delta\phi$.
	\item $\sigma$ is the canonical bundle-morphism
		\begin{displaymath}
			\sigma:  0^\ast \left( V_M W \right)  \; \rightarrow \;  0^\ast \left(V_E W \right) \simeq W
		\end{displaymath}
		 where $V_M W$ and $V_F W$ are the vertical sub-bundle of $T W$ w.r.t. $M$ and $F$ respectively 
		 and $0$ is the zero-section of $W$ over $F$.
\end{itemize}

@@

\begin{remark}
	The Jacobi operator $\left(J(\phi) \right) [\delta \phi]$ around a solution $\phi$ when acting on $\delta \phi \in T_\phi \Conf$ yields a twisted form that can be evaluated on a $V \in \Gamma\left(\phi^\ast(V^\TwistedLinear F )\right)$
\end{remark}

@@

\begin{figure}
\centering
\includegraphics[width=0.40\textwidth]{Pictures/sigma.png}
  %\caption{.}
\end{figure}
Visually this is a map which "verticalize" w.r.t. $E$ the vertical vectors w.r.t. $M$ along $0(E) \subset W$

@@

\newpage
\section{Covariant approach: pre-symplectic form}
\begin{figure}
\centering
\includegraphics[width=0.40\textwidth]{Pictures/symform.png}
  \caption{(temp)}
\end{figure}

@@

Rigourosly, we define $\forall \phi \in \Conf$, $\forall \Sigma \subset M$ $m-1$ dimensional, compact submanifold of $M$:

\begin{definition}[(covariant) pre-symplectic form in $\phi$ pertaining to $\Sigma$]\label{Def:presymform}
	\begin{displaymath}
		\Omega_\Gamma ( \phi ) : \textrm{T}_\phi \Conf \times \textrm{T}_\phi \Conf \rightarrow \Real
\end{displaymath}		
	S.t:
	\begin{displaymath}
		\left\langle	\Omega_\Sigma ( \phi) \right.\left\lvert \delta \phi_1 , \delta \phi_2 \right\rangle =		
		\int_\Sigma (\phi)^\ast \left[ i_{\delta\phi_2} i_{\delta \phi_1} \omega \right] =
		\int_\Sigma \Upsilon_\phi \left( \delta \phi_1 , \delta \phi_2 \right) = 
		\begin{cases}
			\int_\Sigma (j^1 \varphi)^\ast \left[ i_{j^1 \delta\varphi_2} i_{j^1 \delta \varphi_1} \omega_\Lagrangian \right]
			& \qquad \textrm{(Lagrangian)} \\
			\int_\Sigma (\varphi, \pi)^\ast \left[ i_{( \delta \varphi_2, \delta \pi_2)} i_{( \delta \varphi_1, \delta \pi_1)} \omega_\Hamiltonian \right]
			& \qquad \textrm{(Hamiltonian)} 		
		\end{cases}
	\end{displaymath}
where the form $\Upsilon_\phi \left( \delta \phi_1 , \delta \phi_2 \right)  \in \Omega^{m-1}(M)$ to be integred on $\Sigma$ is called \emph{Symplectic current}.
\end{definition}
\begin{notation}
	The usual notation is still employed:
	\begin{displaymath}
		\Upsilon_\phi \left( \delta \phi_1 , \delta \phi_2 \right) = 
			\begin{cases}
				(j^1 \varphi)^\ast \left[ i_{j^1 \delta\varphi_2} i_{j^1 \delta \varphi_1} \omega_\Lagrangian \right]
				& \qquad \textrm{(Lagrangian framework)} \\
				(\varphi, \pi)^\ast \left[ i_{( \delta \varphi_2, \delta \pi_2)} i_{( \delta \varphi_1, \delta \pi_1)} \omega_\Hamiltonian \right]
				& \qquad \textrm{(Hamiltonian framework)} 		
			\end{cases}	
	\end{displaymath}

\end{notation}

The form $\Omega_\Sigma (\phi)$ is manifestly bilinear and skew-symmetric but at this stage is far from obvious to prove exactness and non-degeneracy. We can not conclude yet that it is a symplectic form.
However, the clumsy dependence from the choice $\Sigma$ represent a more urgent problem.
An important remark can be done:

\begin{proposition}\label{Prop:CurrentClosed}
	When restricted to $\textrm{T}_\phi \Sol$, the symplectic current is a closed form. That is:
	\begin{displaymath}
		\textrm{d} \Upsilon_\phi \left( \delta \phi_1 , \delta \phi_2 \right) = 0 \qquad \delta \phi_1 , \delta \phi_2 \in T_\phi \Sol
	\end{displaymath}
\end{proposition}
\begin{proof}
	Working in the collective notation introduced above, we have:
	\begin{align*}
		\textrm{d} \Upsilon_\phi \left( \delta \phi_1 , \delta \phi_2 \right) &= 
		\phi^\ast \textrm{d} \left[ i_{\delta\phi_2} i_{\delta \phi_1} \omega \right] = 
		\phi^\ast \left( - i_{\delta\phi_2} \textrm{d} ( i_{\delta \phi_1} \omega )  + \Lie_{\delta\phi_2 } ( i_{\delta \phi_1} \omega ) \right)
	\end{align*}

\end{proof}

This nice condition represent a sort of exactness condition for the covariant pre-symplectic form.\\
However, it is not sufficient to eliminate the unwanted dependence from the choice of $\Sigma$.
In order to get rid of this drawback are to be introduced some other assumptions.\\
It is customary to require the working condition of physical field system theories on curved background:
\begin{enumerate}
	\item $M$ is a \emph{globally-hyperbolic spacetime}
	\item $\Sigma$ is an arbitrary cauchy surface
	\item $T_\phi \Conf =\Gamma_{sc} (V^\TwistedLinear F )$ i.e. are only considered space-like compact variations.
\end{enumerate}
The first condition guarantees that exists a Cauchy surface and that the spacetime can be folieted according to it $M= \Sigma \times \Real$.\\
Considering a spacetime \footnote{Oriented, time-oriented, connected, Lorentzian smooth manifold. } allow us to define a space-like compact support.\\
From last condition follows that for any Cauchy surface $\Sigma$, $ \supp(\Upsilon_\phi (\delta \phi_1, \delta \phi_2) \cap \Gamma$ is compact, thus the integration involved in definition \ref{Def:presymform} is always well-defined.

\begin{proposition}
	In this condition $\Omega_\Sigma (\phi): T_\phi \Sol \times T_\phi \Sol \rightarrow \Real$ does not depends on the choice of $\Sigma$ anymore.
\end{proposition}
\begin{proof}
	Consider 
	\begin{displaymath}
		K_t = \supp\left(\Upsilon_\phi (\delta \phi_1, \delta \phi_2) \right) \cap \Gamma_t
	\end{displaymath}
	by varying $t$, it determines an m-dimensional domain $V \subset M$.
	For any $U$ open set in $M$ containing $K={K_t}_{0<t<t'}$ we have, as a simple consequence of proposition \ref{Prop:CurrentClosed} and Stokes' lemma, that
	\begin{displaymath}
		\int_V \textrm{d} \Upsilon_\phi = - \int_{\Gamma_0} \Upsilon_\phi +\int_{\Gamma_t} \Upsilon_\phi 
	\end{displaymath}
	Therefore $\Omega_\Sigma (\phi) = \Omega_{\Sigma'} (\phi)$ for all possible pair of Cauchy surfaces.
\end{proof}

\begin{wrapfigure}{L}{0.41\textwidth}
\centering
\includegraphics[width=0.40\textwidth]{Pictures/sigma.png}
  %\caption{.}
\end{wrapfigure}

### 2017-05-17 ms_sigmaconstruction.tex
Almost finished..

\begin{tikzpicture}[node distance=3.25cm,
    every node/.style={fill=white, font=\sffamily}, align=center]

@@

  \node (pbTW_block) [ tangent, above right of= TW_block, xshift=0.4cm , label={[label distance=-0.775cm]90:$\pi_{E,W}^\ast\left( TE\right)$}]	{};
  \node (pbVmW_block)	[verticalbig, above right of= TW_block , xshift=0.4cm, yshift=-0.40cm]	{$\pi_{E,W}^\ast\left( TV\right)$};

  \end{tikzpicture}

  \begin{tikzpicture}
    \node[matrix] (A) {
        \draw (0,0) rectangle (1,1); 
        \node at (0.5,0.5) {A}; \\
    };
    \node[matrix,left of=A] (B) 
    {
        \draw (0.5,0.5) circle (0.5);
        \fill (0.5,0.5) circle (0.1); \\
    };
\end{tikzpicture}
\end{document}

>>>>

\begin{tikzpicture}[node distance=4cm,
    every node/.style={fill=white, font=\sffamily}, align=center]

@@

  \node (pbTW_block) [ tangent, above right of= TW_block, xshift=0.4cm , label={[label distance=-0.9cm]90:$\pi_{E,W}^\ast\left( TE\right)$}]	{};
  \node (pbVmW_block)	[verticalbig, above right of= TW_block , xshift=0.4cm, yshift=-0.35cm]	{$\pi_{E,W}^\ast\left( TV\right)$};

  % Bundle Projections 
	\draw[projection]	(E_block) -- (M_block);
	\draw[projection]	([xshift=-0.5cm]W_block.south) -- node {$\pi$} ([xshift=-0.5cm]E_block.north);
	\draw[projection]	(TE_block) -- (E_block);
	\draw[projection]	(TW_block) -- (W_block);
	\draw[projection]	(pbTW_block.west) -- (W_block.north);

	\draw[lagmap]	(pbTW_block.south) -- node {$m$}	(TE_block.east);   
	\draw[projection]	([xshift=-0.5cm]TW_block.south) -- node {$T \pi$}	([xshift=-0.5cm]TE_block.north); 
	\draw[lagmap]	([xshift=+0.5cm] E_block.north) -- node {$0$}	([xshift=+0.5cm] W_block.south);    
	\draw[lagmap]	([xshift=+0.5cm] TE_block.north) -- node {$T 0$}	([xshift=+0.5cm] TW_block.south);   	
	\draw[lagmap]	(pbTW_block) -- node {$T \pi$}	(TW_block);

  \end{tikzpicture}
\end{document}
